---
title: "p8105_hw1_sl4655"
author: "Shuya Liu"
date: "September 17, 2019"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

#load packages
library(tidyverse)
```

# _Problem 0.1_
This “problem” focuses on the use of R Markdown to write reproducible reports, GitHub for version control, and R Projects to organize your work.

To that end:

* create a public GitHub repo + local R Project; we suggest naming this repo / directory p8105_hw1_YOURUNI (e.g. p8105_hw1_ajg2202 for Jeff), but that’s not required
* create a single .Rmd file named p8105_hw1_YOURUNI.Rmd that renders to github_document
* submit a link to your repo via Courseworks

Your solutions to Problems 1 and 2 should be implemented in your .Rmd file, and your git commit history should reflect the process you used to solve these Problems.

# _Problem 0.2_
This “problem” focuses on correct styling for your solutions to Problems 1 and 2. We will look for:

* meaningful variable / object names
* readable code (one command per line; adequate whitespace and indentation; etc)
* clearly-written text to explain code and results
* a lack of superfluous code (no unused variables are defined; no extra library calls; ect)

# _Problem 1_
This problem is intended to emphasize variable types and introduce coercion; some awareness of how R treats numeric, character, and factor variables is necessary for working with these data types in practice.

## Create a data frame

Create a data frame comprised of:

* a random sample of size 8 from a standard Normal distribution
* a logical vector indicating whether elements of the sample are greater than 0
* a character vector of length 8
* a factor vector of length 8, with 3 different factor “levels”

Try to take the mean of each variable in your dataframe. What works and what doesn’t?

```{r data_frame_1, warning = FALSE, collapse = TRUE}
# Create an original data frame
df_1 = tibble(
  vec_numeric = rnorm(8, mean = 0, sd = 1),
  vec_logical = vec_numeric > 0,
  vec_character = c("James", "Benedict","Kate","Jennifer","Jake","Morgan","Martin","Natalie"),
  vec_factor = factor(c("B","A","C","B","B","A","C","C"))
)

# Take the mean of each variable
mean(pull(df_1, vec_numeric))
mean(pull(df_1, vec_logical))
mean(pull(df_1, vec_character))
mean(pull(df_1, vec_factor))
```
* **_Some Observations_**: As we can see from above, by taking the mean of the numeric and logical vectors, we get `r round(mean(pull(df_1, vec_numeric)), 3)` 
                    and `r round(mean(pull(df_1, vec_logical)), 3)` respectively. 
But when it comes to the character vector and the factor, we only get `NA`, because the `mean()` function only works on numeric and logical vectors, otherwise we will get a warning message `argument is not numeric or logical: returning NA`.

## Coersion as.numeric

In some cases, you can explicitly convert variables from one type to another. Write a code chunk that applies the `as.numeric` function to the logical, character, and factor variables (please show this chunk but not the output). What happens, and why? Does this help explain what happens when you try to take the mean?
```{r numeric, eval = FALSE}
as.numeric(pull(df_1, vec_logical))
as.numeric(pull(df_1, vec_character))
as.numeric(pull(df_1, vec_factor))
```
* **_Some Observations_**: 
    * For logical variables, the `as.numeric` function coerses the _logical variables_ into _numeric variables_. Specifically, `TRUE` equals 1 and `FALSE` equals 0
    * For character variables, the outcome is `NA` and a warning shows `NAs introduced by coercion`, which means that character variables cannot be coersed into numeric ones. That would explain why the `mean` of the charater variables is `NA`
    * For factor variables, the coersion turns them into their levels, and autmatically assigns the factors with levels from 1 to 3. But, it doesn't explain why the `mean` was `NA` intuitively

## Convert logical vector

In a second code chunk:

* convert the logical vector to numeric, and multiply the random sample by the result
* convert the logical vector to a factor, and multiply the random sample by the result
* convert the logical vector to a factor and then convert the result to numeric, and multiply the random sample by the result

```{r coercion, collapse = TRUE}
# Convert the logical vector to numeric
log_num = as.numeric(pull(df_1, vec_logical))
log_num * pull(df_1, vec_numeric)

# Covert the logical vector to a factor
log_fac = factor(pull(df_1, vec_logical))
log_fac * pull(df_1, vec_numeric)

# Convert the logical vector to a factor and then convert the result to numeric
log_fac_num = as.numeric(log_fac)
log_fac_num * pull(df_1, vec_numeric)
```


# _Problem 2_
This problem focuses the use of inline R code, plotting, and the behavior of `ggplot` for variables of different types.

* Create a data frame comprised of:
    * `x`: a random sample of size 500 from a standard Normal distribution
    * `y`: a random sample of size 500 from a standard Normal distribution
    * A logical vector indicating whether `x + y > 1`
    * A numeric vector created by coercing the above logical vector
    * A factor vector created by coercing the above logical vector
    
* Write a short description of your vector using inline R code, including: 
    * the size of the dataset (using `nrow` and `ncol`) 
    * the mean, median, and standard deviation of `x` 
    * the proportion of cases for which `x + y > 1`

## Create a data frame    
```{r data_frame_2}
df_2 = tibble(
  x = rnorm(n = 500, mean = 0, sd = 1),
  y = rnorm(n = 500, mean = 0, sd = 1),
  vec_logical = (x + y) > 1,
  vec_numeric = as.numeric(vec_logical),
  vec_factor = factor(vec_logical)
)
```
**_Description_**: 

* The dataset has `r nrow(df_2)` rows and `r ncol(df_2)` columns
* For `x`, 
    * the mean is `r round(mean(pull(df_2, x)), 3)`
    * the median is `r round(median(pull(df_2, x)), 3)`
    * the standard deviation is `r round(sd(pull(df_2, x)), 3)`
* The proportion of cases for which `x + y > 1` is `r sum(pull(df_2, vec_logical)) / nrow(df_2)`

## Scatterplots

Make a scatterplot of y vs x; color points using the logical variable (adding color = ... inside of aes in your ggplot code should help). Make a second and third scatterplot that color points using the numeric and factor variables, respectively, and comment on the color scales.

* Scatterplot 1
```{r scatterplot_1}
# Using the logical variable
ggplot(df_2, aes(x = x, y = y)) + 
  geom_point(aes(color = vec_logical))
```

* Scatterplot 2
```{r scatterplot_2}
# Using the numeric variable
ggplot(df_2, aes(x = x, y = y)) + 
  geom_point(aes(color = vec_numeric))
```

* Scatterplot 3
```{r scatterplot_3}
# Using the factor variable
ggplot(df_2, aes(x = x, y = y)) + 
  geom_point(aes(color = vec_factor))
```

