---
title: "p8105_hw1_sl4655"
author: "Shuya Liu"
date: "September 17, 2019"
output: github_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

#load packages
library(tidyverse)
```


# _Problem 1_
This problem is intended to emphasize variable types and introduce coercion; some awareness of how R treats numeric, character, and factor variables is necessary for working with these data types in practice.


### Create a data frame

* a random sample of size 8 from a standard Normal distribution
* a logical vector indicating whether elements of the sample are greater than 0
* a character vector of length 8
* a factor vector of length 8, with 3 different factor “levels”

```{r data_frame_1}
# Create an original data frame
df_1 = tibble(
  vec_numeric = rnorm(8, mean = 0, sd = 1),
  vec_logical = vec_numeric > 0,
  vec_character = c("James", "Benedict","Kate","Jennifer","Jake","Morgan","Martin","Natalie"),
  vec_factor = factor(c("B","A","C","B","B","A","C","C"))
)
```

##### Try to take the mean of each variable in your dataframe. What works and what doesn’t?

```{r mean, warning = FALSE, collapse = TRUE}
# Take the mean of each variable
mean(pull(df_1, vec_numeric))
mean(pull(df_1, vec_logical))
mean(pull(df_1, vec_character))
mean(pull(df_1, vec_factor))
```

* As we can see from above, 
    * the `mean()` function works for numeric and logic value (taking true as 1 and false as 0),
    * it doesn't work on character vector and the factor, only returning `NA`.
        * The reason is that the `mean()` function only works on numeric and logical vectors, and in this case the factors are in characters, so R make the factor and character vectors into a numeric vector

### Coersion as.numeric

In some cases, you can explicitly convert variables from one type to another. Write a code chunk that applies the `as.numeric` function to the logical, character, and factor variables (please show this chunk but not the output). 
```{r numeric, eval = FALSE}
as.numeric(pull(df_1, vec_logical))
as.numeric(pull(df_1, vec_character))
as.numeric(pull(df_1, vec_factor))
```

##### What happens, and why? Does this help explain what happens when you try to take the mean?

* For logical variables, the `as.numeric` function coerses the _logical variables_ into _numeric variables_. Specifically, `TRUE` equals 1 and `FALSE` equals 0
* For character variables, the outcome is `NA` and a warning shows `NAs introduced by coercion`, which means that character variables cannot be coersed into numeric ones. That would explain why the `mean` of the character variables is `NA`
* For factor variables, the coersion turns them into their levels, and autmatically assigns the factors with levels from 1 to 3.

### Convert logical vector

* convert the logical vector to numeric, and multiply the random sample by the result
* convert the logical vector to a factor, and multiply the random sample by the result
* convert the logical vector to a factor and then convert the result to numeric, and multiply the random sample by the result

```{r coercion, collapse = TRUE}
# Convert the logical vector to numeric
log_num = as.numeric(pull(df_1, vec_logical))
log_num * pull(df_1, vec_numeric)

# Covert the logical vector to a factor
log_fac = factor(pull(df_1, vec_logical))
log_fac * pull(df_1, vec_numeric)

# Convert the logical vector to a factor and then convert the result to numeric
log_fac_num = as.numeric(log_fac)
log_fac_num * pull(df_1, vec_numeric)
```
* logical to numeric: returns numerical values, because `TRUE` equals 1, and `FAlSE` equals 0
* factor to character to numeric: returns NAs, because when factors are turned into characters, the as.numeric function will return NA for characters
* character to factor to numeric: returns numerical value, because characters can turned into factors, and then as.numeric assign numbers to each level of the factor


# _Problem 2_
This problem focuses the use of inline R code, plotting, and the behavior of `ggplot` for variables of different types.

* Create a data frame comprised of:
    * `x`: a random sample of size 500 from a standard Normal distribution
    * `y`: a random sample of size 500 from a standard Normal distribution
    * A logical vector indicating whether `x + y > 1`
    * A numeric vector created by coercing the above logical vector
    * A factor vector created by coercing the above logical vector
    
* Write a short description of your vector using inline R code, including: 
    * the size of the dataset (using `nrow` and `ncol`) 
    * the mean, median, and standard deviation of `x` 
    * the proportion of cases for which `x + y > 1`

### Create a data frame    
```{r data_frame_2}
set.seed(50) # set.seed() would fix the randomly chosen variables

df_2 = tibble(
  x = rnorm(n = 500, mean = 0, sd = 1),
  y = rnorm(n = 500, mean = 0, sd = 1),
  vec_logical = (x + y) > 1,
  vec_numeric = as.numeric(vec_logical),
  vec_factor = factor(vec_logical)
)
```

* The dataset has `r nrow(df_2)` rows and `r ncol(df_2)` columns
* For `x`, 
    * the mean is `r round(mean(pull(df_2, x)), 3)`
    * the median is `r round(median(pull(df_2, x)), 3)`
    * the standard deviation is `r round(sd(pull(df_2, x)), 3)`
* The proportion of cases for which `x + y > 1` is `r sum(pull(df_2, vec_logical)) / nrow(df_2)`

### Scatterplots

Make a scatterplot of y vs x; color points using the logical variable (adding color = ... inside of aes in your ggplot code should help). Make a second and third scatterplot that color points using the numeric and factor variables, respectively, and comment on the color scales.

##### Scatterplot 1
```{r scatterplot_1}
# Using the logical variable as determinants for different colors
plot_1 = ggplot(df_2, aes(x = x, y = y, color = vec_logical)) + 
  geom_point() + 
  ggtitle("Color by Logical")
# Using the numeric variable as determinants for different colors
plot_2 = ggplot(df_2, aes(x = x, y = y, color = vec_numeric)) + 
  geom_point() + 
  ggtitle("Color by Numeric")
# Using the factor variable as determinants for different colors
plot_3 = ggplot(df_2, aes(x = x, y = y, color = vec_factor)) + 
  geom_point() + 
  ggtitle("Color by Factor")

#Show the plots
plot_1
plot_2
plot_3
```


##### Export the first scatterplot
```{r export}
ggsave("Plot Color by Logical.png", plot = plot_1)
```